---
title: "buffer_gbif"
author: "Zoë Hillier-Weltman"
date: "08/08/2022"
output: html_document
---
```{r}
library(tidyverse)
library(sf)
library(sp)
library(rgdal)
library(raster)
library(rgeos)
library(here)
```

sites being used: 31
carrizo 1-4, carrizo 6, avenal, semitropic, antelope valley- don't have site file, coalinga, silver creek, tejon, cuyama 1-6, panoche 1-3, lokern, mojave 1-5, tecopa, barstow 1-2, HoM, sheephole

recursive
```{r}
data_path <- "sites"

read_zip = function(f){
 s = sf::st_read(f)
 return(s[,c("Name")])
}
polygons_read <- as.list(dir(data_path, pattern="*.kml", full.names=T)) 
polygons<-do.call(rbind, lapply(polygons_read, read_zip))

#changing format
polygon_wkts<-st_as_text(polygons$geometry)

mean(st_area(polygons))#10km^2 should be bigger than 0.27km^2

```

or prog report- 30 initial sites (21 w shrubs)
```{r}
prog_polys<-polygons%>%
  dplyr::filter(Name %in% (c("Avenal","Barstow_1","Barstow_2","Carrizo_1","Carrizo_2","Carrizo_3","Carrizo_4","Carrizo_6","Coalinga","Cuyama_1","Cuyama_2","Cuyama_3","Cuyama_4","Cuyama_5","Cuyama_6","HeartofMojave","Lokern","Mojave_1","Mojave_2","Mojave_3","Mojave_4","Mojave_5","Panoche_1","Panoche_2","Panoche_3","Panoche_silvercreek","Semitropic","Sheephole","Tejon_ERG","Tecopa"
                )))
```

figure out what area of og sites was, because the buffered centroids look small atm
```{r}
polygons$area <- st_area(polygons) #Take care of units
```

```{r}
library(wellknown)
library(spocc)
library(buffeRs)
points_centroids<-st_centroid(prog_polys)

x = 4326 #crs for first spatial df
#10km^2 for area for buffered sites

#data, read in, set as sf, list projection####
centroids <- points_centroids %>% 
  st_as_sf(coords=c("long","lat"), crs = x, remove=FALSE) %>% 
  st_transform(32610)%>%
  st_make_valid()#crs = 32610 is UTM and will return distances in meters

```

copy-paste the 30 buffered sites since iteration not yet working
```{r}
ave = sf::st_point(c(761052.797999445, 3994672.67159508))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

st_area(ave)

bar1 = sf::st_point(c(1062342.88500539, 3900746.85342364))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

st_area(carr)

bar2 = sf::st_point(c(958163.878846229, 3886789.60811268))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

car1 = sf::st_point(c(798199.200708459, 3900375.67244386))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

car2 = sf::st_point(c(799332.001523994, 3899358.7450804))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

car3 = sf::st_point(c(802755.396426991, 3896349.55239856))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

car4 = sf::st_point(c(807920.807191618, 3891248.71057087))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)
  
car6 = sf::st_point(c(791819.390072172, 3904507.1129823))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

coal = sf::st_point(c(742418.340269413, 4010965.26229504))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy1 = sf::st_point(c(821512.725205021, 3861921.83967465))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy2 = sf::st_point(c(821212.832738582, 3862400.31508178))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy3 = sf::st_point(c(821443.940153515, 3871872.32605763))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy4 = sf::st_point(c(821485.125033712, 3872838.75317886))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy5 = sf::st_point(c(783763.411643091, 3882551.09282527))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy6 = sf::st_point(c(783133.59718452, 3882887.77652211))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

hom = sf::st_point(c(1170687.69075316, 3863937.88019742))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

lok = sf::st_point(c(810421.963184896, 3917631.71201444))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

moj1 = sf::st_point(c(1132999.20638878, 3923622.19822479))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

moj2 = sf::st_point(c(1144423.11243581, 3923725.60265459))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

moj3 = sf::st_point(c(1150560.42677661, 3916827.91002961))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

moj4 = sf::st_point(c(1157281.89900324, 3913999.96584649))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

moj5 = sf::st_point(c(1162950.89096981, 3913332.01494015))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

pan1 = sf::st_point(c(696509.188076656, 4063725.48583983))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

pan2 = sf::st_point(c(695891.752914405, 4064372.25008394))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

pan3 = sf::st_point(c(695763.670426077, 4064156.55005868))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

pansl = sf::st_point(c(707039.464984836, 4051533.2702302))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

semi = sf::st_point(c(806660.672891432, 3951300.60690333))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

sheep= sf::st_point(c(1171450.01712319, 3809018.69261025))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

tec = sf::st_point(c(1115892.95183313, 3989096.65725961))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

plot(tec)

tej = sf::st_point(c(902033.825842286, 3868120.02721386))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(3200,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)
  
st_area(tec)

buffered_wkts_prog<-rbind(ave,bar1,bar2,car1,car2,car3,car4,car6,coal,cuy1,cuy2,cuy3,cuy4,cuy5,cuy6,hom,lok,moj1,moj2,moj3,moj4,moj5,pan1,pan2,pan3,pansl,semi,sheep,tec,tej)

```

now gbif pull for 30 buffered sites
```{r}
library(rgbif)
library(spocc)

user<- "zoehw2"
email <- "zoe.hillier.weltman@gmail.com"
pwd <- "foxhunter"

#aves taxon key is 212

#not sure if there's a better workaround
for (i in buffered_wkts_prog[1:3]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[4:6]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[7:9]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[10:12]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[13:15]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[16:18]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[19:21]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[22:24]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[25:27]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[28:30]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

#double checking barstow and coalinga

for (i in buffered_wkts_prog[2]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[9]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

```

and now for the buffered sites
```{r}
data_path_3 <- "gbif_sites_square"

read_csv = function(f){
 s = read_csv(f)
 return()
}

#gbif_read <- as.list(dir(data_path_3, pattern="*.csv", full.names=T)) 
#gbif_buffered<-do.call(rbind, lapply(gbif_read, read_csv))

library(readr)
Avenal <- read_delim("gbif_sites_square/avenal.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Avenal$site<-"Avenal"

Barstow_2 <- read_delim("gbif_sites_square/bar2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Barstow_2$site<-"Barstow_2"

Carrizo_1 <- read_delim("gbif_sites_square/car1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_1$site<-"Carrizo_1"

Carrizo_2 <- read_delim("gbif_sites_square/car2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_2$site<-"Carrizo_2"

Carrizo_3 <- read_delim("gbif_sites_square/car3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_3$site<-"Carrizo_3"

Carrizo_4 <- read_delim("gbif_sites_square/car4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_4$site<-"Carrizo_4"

Carrizo_6 <- read_delim("gbif_sites_square/car6.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_6$site<-"Carrizo_6"

Cuyama_1 <- read_delim("gbif_sites_square/cuy1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_1$site<-"Cuyama_1"

Cuyama_2 <- read_delim("gbif_sites_square/cuy2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_2$site<-"Cuyama_2"

Cuyama_3 <- read_delim("gbif_sites_square/cuy3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_3$site<-"Cuyama_3"

Cuyama_4 <- read_delim("gbif_sites_square/cuy4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_4$site<-"Cuyama_4"

Cuyama_5 <- read_delim("gbif_sites_square/cuy5.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_5$site<-"Cuyama_5"

Cuyama_6 <- read_delim("gbif_sites_square/cuy6.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_6$site<-"Cuyama_6"

Heartofmojave <- read_delim("gbif_sites_square/hom.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Heartofmojave$site<-"Heartofmojave"

Lokern <- read_delim("gbif_sites_square/lok.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Lokern$site<-"Lokern"

Mojave_1 <- read_delim("gbif_sites_square/mjv1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_1$site<-"Mojave_1"

Mojave_2 <- read_delim("gbif_sites_square/mjv2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_2$site<-"Mojave_2"

Mojave_3 <- read_delim("gbif_sites_square/mjv3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_3$site<-"Mojave_3"

Mojave_4 <- read_delim("gbif_sites_square/mjv4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_4$site<-"Mojave_4"

Mojave_5 <- read_delim("gbif_sites_square/mjv5.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_5$site<-"Mojave_5"

Panoche_1 <- read_delim("gbif_sites_square/pan1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_1$site<-"Panoche_1"

Panoche_2 <- read_delim("gbif_sites_square/pan2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_2$site<-"Panoche_2"

Panoche_3 <- read_delim("gbif_sites_square/pan3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_3$site<-"Panoche_3"

Panoche_silvercreek <- read_delim("gbif_sites_square/pansl.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_silvercreek$site<-"Panoche_silvercreek"

Semitropic <- read_delim("gbif_sites_square/semitropic.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Semitropic$site<-"Semitropic"

Sheephole <- read_delim("gbif_sites_square/sheephole.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Sheephole$site<-"Sheephole"

Tejon_ERG <- read_delim("gbif_sites_square/tejon.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Tejon_ERG$site<-"Tejon_ERG"

Tecopa <- read_delim("gbif_sites_square/tecopa.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Tecopa$site<-"Tecopa"

gbif_sites_buffered<-rbind(Avenal,Barstow_2,Carrizo_1,Carrizo_2,Carrizo_3,Carrizo_4,Carrizo_6,Cuyama_1,Cuyama_2,Cuyama_3,Cuyama_4,Cuyama_5,Cuyama_6,Heartofmojave,Lokern,Mojave_1,Mojave_2,Mojave_3,Mojave_4,Mojave_5,Panoche_1,Panoche_2,Panoche_3,Panoche_silvercreek,Semitropic,Sheephole,Tecopa,Tejon_ERG)
```

filter out nas, create new rows for individualcount>1
```{r}
#if individual count = na
gbif_sites_buffered<-gbif_sites_buffered %>% drop_na(individualCount)

#gbif_buffered<-gbif_sites_buffered%>%distinct(gbifID, .keep_all = TRUE)

expanded_sites<- data.frame(gbif_buffered_uniques[rep(seq_len(dim(gbif_buffered_uniques)[1]),  with(gbif_buffered_uniques, ifelse(individualCount > 0 & !is.na(individualCount), individualCount, 1))
), , drop = FALSE], row.names=NULL)

expanded_buffered_uniques$individualCount<-1

expanded_buffered_uniques<-rename(expanded_buffered_uniques,lat=decimalLatitude)
expanded_buffered_uniques<-rename(expanded_buffered_uniques,long=decimalLongitude)
expanded_buffered_uniques<-rename(expanded_buffered_uniques,site_code=site)
expanded_buffered_uniques$rep<-1:nrow(expanded_buffered_uniques)
expanded_buffered_uniques <- expanded_buffered_uniques[colnames(expanded_buffered_uniques)[c(52,1:51)]]

#if individual count >1
expanded_sites <- data.frame(gbif_sites_buffered[rep(seq_len(dim(gbif_sites_buffered)[1]),  with(gbif_sites_buffered, ifelse(individualCount > 0 & !is.na(individualCount), individualCount, 1))
), , drop = FALSE], row.names=NULL)

expanded_sites$individualCount<-1
expanded_sites$rep<-1:nrow(expanded_sites)
expanded_sites <- expanded_sites[colnames(expanded_sites)[c(52,1:51)]]

expanded_sites<-rename(expanded_sites,lat=decimalLatitude)
expanded_sites<-rename(expanded_sites,long=decimalLongitude)
expanded_sites<-rename(expanded_sites,site_code=site)
#there are duplicates due to overlap

#renaming site codes to match regional file

site_names<-expanded_sites%>%
  group_by(site_code)%>%
  summarise()


#write.csv(expanded_sites, file = "buffer_gbif.csv")

write.csv(expanded_sites, file = "buffer_gbif.csv")
saveRDS(expanded_sites, file = "buffered_site_gbif.rds")
readRDS("buffered_site_gbif.rds")
```