---
title: "buffer_gbif"
author: "Zoë Hillier-Weltman"
date: "08/08/2022"
output: html_document
---
```{r}
library(tidyverse)
library(sf)
library(sp)
library(rgdal)
library(raster)
library(rgeos)
library(here)
```

sites being used: 31
carrizo 1-4, carrizo 6, avenal, semitropic, antelope valley- don't have site file, coalinga, silver creek, tejon, cuyama 1-6, panoche 1-3, lokern, mojave 1-5, tecopa, barstow 1-2, HoM, sheephole

recursive
```{r}
data_path <- "sites"

read_zip = function(f){
 s = sf::st_read(f)
 return(s[,c("Name")])
}
polygons_read <- as.list(dir(data_path, pattern="*.kml", full.names=T)) 
polygons<-do.call(rbind, lapply(polygons_read, read_zip))

#changing format
polygon_wkts<-st_as_text(polygons$geometry)

mean(st_area(polygons))#10km^2 should be bigger than 0.27km^2

```

or prog report- 30 initial sites (21 w shrubs)
```{r}
prog_polys<-polygons%>%
  dplyr::filter(Name %in% (c("Avenal","Barstow_1","Barstow_2","Carrizo_1","Carrizo_2","Carrizo_3","Carrizo_4","Carrizo_6","Coalinga","Cuyama_1","Cuyama_2","Cuyama_3","Cuyama_4","Cuyama_5","Cuyama_6","HeartofMojave","Lokern","Mojave_1","Mojave_2","Mojave_3","Mojave_4","Mojave_5","Panoche_1","Panoche_2","Panoche_3","Panoche_silvercreek","Semitropic","Sheephole","Tejon_ERG","Tecopa"
                )))
```

figure out what area of og sites was, because the buffered centroids look small atm
```{r}
polygons$area <- st_area(polygons) #Take care of units
```

```{r}
library(wellknown)
library(spocc)
points_centroids<-st_centroid(prog_polys)

crsuggest::suggest_crs(polygons, type = "projected", units='m')

x = 4326 #crs for first spatial df
y = 4326 #crs for second spatial df
z = 1800 #set buffer is using a join to point data 
#10km^2 for area for buffered sites

#data, read in, set as sf, list projection####
centroids <- points_centroids %>% 
  st_as_sf(coords=c("long","lat"), crs = x, remove=FALSE) %>% 
  st_transform(32610)%>%
  st_make_valid()#crs = 32610 is UTM and will return distances in meters

#create buffer for point data

buffered_centroids<-st_buffer(centroids, z)%>%
  st_transform(4326)%>%
  st_make_valid()

st_is_valid(buffered_centroids)

st_geometry_type(buffered_centroids)

st_coordinates(buffered_centroids)

st_area(buffered_centroids)

plot(buffered_centroids)

st_bbox(buffered_centroids)#a way to run for each element of the list?

bboxes_buffered_sites = cbind(st_sf(geometry=buffered_centroids$geometry),do.call(rbind,lapply(buffered_centroids$geometry, st_bbox)))

centroid_wkts<-st_as_text(buffered_centroids$geometry)#yielding too many points? invalid shape for gbif
```

copy-paste the 30 buffered sites since iteration not yet working
```{r}
ave<-bbox2wkt(
-120.1216,
36.04542,
-120.0817,
36.07784)

bar1<-bbox2wkt(-116.8542,
35.07644,
-116.8149,
35.10877)
  
bar2<-bbox2wkt(
-117.9998,
35.00375,
-117.9604,
35.03613)

car1<-bbox2wkt(-119.7444,
35.18592,
-119.7049,
35.21834)
  
car2<-bbox2wkt(-119.7323,
35.17643,
-119.6928,
35.20885)
  
car3<-bbox2wkt(-119.6959,
35.14832,
-119.6564,
35.18074)
  
car4<-bbox2wkt(-119.6412,
35.10083,
-119.6017,
35.13325)
  
car6<-bbox2wkt(-119.8129,
35.22499,
-119.7734,
35.25741)
  
coal<-bbox2wkt(-120.3233,
36.19697,
-120.2833,
36.22939)

cuy1<-bbox2wkt(-119.5035,
34.83263,
-119.4642,
34.86505)

cuy2<-bbox2wkt(-119.5066,
34.83704,
-119.4673,
34.86945)

cuy3<-bbox2wkt(-119.5005,
34.92222,
-119.4611,
34.95463)

cuy4<-bbox2wkt(-119.4997,
34.93090,
-119.4603,
34.96332)

cuy5<-bbox2wkt(-119.9088,
35.02961,
-119.8694,
35.06204)

cuy6<-bbox2wkt(-119.9156,
35.03282,
-119.8762,
35.06525)

hom<-bbox2wkt(-115.7036,
34.68124,
-115.6645,
34.71352)

lok<-bbox2wkt(-119.6039,
35.33754,
-119.5643,
35.36996)

mj1<-bbox2wkt(-116.0669,
35.23945,
-116.0275,
35.27174)

mj2<-bbox2wkt(-115.9421,
35.23312,
-115.9028,
35.26541)

mj3<-bbox2wkt(-115.8806,
35.16745,
-115.8413,
35.19974)

mj4<-bbox2wkt(-115.8096,
35.13779,
-115.7703,
35.17007)

mj5<-bbox2wkt(-115.7484,
35.12811,
-115.7091,
35.16039)

pan1<-bbox2wkt(-120.8204,
36.68268,
-120.7801,
36.71511)

pan2<-bbox2wkt(-120.8271,
36.68863,
-120.7868,
36.72106)

pan3<-bbox2wkt(-120.8286,
36.68672,
-120.7883,
36.71915)

pansl<-bbox2wkt(-120.7058,
36.57062,
-120.6656,
36.60305)

semi<-bbox2wkt(-119.6326,
35.64176,
-119.5928,
35.67417)

sh<-bbox2wkt(-115.7382,
34.18949,
-115.6993,
34.22178)

tec<-bbox2wkt(-116.2040,
35.83628,
-116.1644,
35.86859)

tej<-bbox2wkt(-118.6220,
34.85976,
-118.5827,
34.89216)

buffered_wkts_prog<-rbind(ave,bar1,bar2,car1,car2,car3,car4,car6,coal,cuy1,cuy2,cuy3,cuy4,cuy5,cuy6,hom,lok,mj1,mj2,mj3,mj4,mj5,pan1,pan2,pan3,pansl,semi,sh,tec,tej)

```

now gbif pull for 30 buffered sites
```{r}
library(rgbif)
library(spocc)

user<- "zoehw2"
email <- "zoe.hillier.weltman@gmail.com"
pwd <- "foxhunter"

#aves taxon key is 212

#not sure if there's a better workaround
for (i in buffered_wkts_prog[1:3]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[4:6]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[7:9]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[10:12]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[13:15]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[16:18]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[19:21]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[22:24]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[25:27]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in buffered_wkts_prog[28:30]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

```

and now for the buffered sites
```{r}
data_path_3 <- "gbif_sites_buffered"

read_csv = function(f){
 s = read_csv(f)
 return()
}

#gbif_read <- as.list(dir(data_path_3, pattern="*.csv", full.names=T)) 
#gbif_buffered<-do.call(rbind, lapply(gbif_read, read_csv))

library(readr)
Avenal <- read_delim("gbif_sites_buffered/avenal.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Avenal$site<-"Avenal"

Barstow_2 <- read_delim("gbif_sites_buffered/bar2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Barstow_2$site<-"Barstow_2"

Carrizo_1 <- read_delim("gbif_sites_buffered/car1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_1$site<-"Carrizo_1"

Carrizo_2 <- read_delim("gbif_sites_buffered/car2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_2$site<-"Carrizo_2"

Carrizo_3 <- read_delim("gbif_sites_buffered/car3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_3$site<-"Carrizo_3"

Carrizo_4 <- read_delim("gbif_sites_buffered/car4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_4$site<-"Carrizo_4"

Carrizo_6 <- read_delim("gbif_sites_buffered/car6.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_6$site<-"Carrizo_6"

Cuyama_1 <- read_delim("gbif_sites_buffered/cuy1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_1$site<-"Cuyama_1"

Cuyama_2 <- read_delim("gbif_sites_buffered/cuy2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_2$site<-"Cuyama_2"

Cuyama_3 <- read_delim("gbif_sites_buffered/cuy3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_3$site<-"Cuyama_3"

Cuyama_4 <- read_delim("gbif_sites_buffered/cuy4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_4$site<-"Cuyama_4"

Cuyama_5 <- read_delim("gbif_sites_buffered/cuy5.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_5$site<-"Cuyama_5"

Cuyama_6 <- read_delim("gbif_sites_buffered/cuy6.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_6$site<-"Cuyama_6"

HeartofMojave <- read_delim("gbif_sites_buffered/hom.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
HeartofMojave$site<-"HeartofMojave"

Lokern <- read_delim("gbif_sites_buffered/lok.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Lokern$site<-"Lokern"

Mojave_1 <- read_delim("gbif_sites_buffered/mj1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_1$site<-"Mojave_1"

Mojave_2 <- read_delim("gbif_sites_buffered/mj2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_2$site<-"Mojave_2"

Mojave_3 <- read_delim("gbif_sites_buffered/mj3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_3$site<-"Mojave_3"

Mojave_4 <- read_delim("gbif_sites_buffered/mj4.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_4$site<-"Mojave_4"

Mojave_5 <- read_delim("gbif_sites_buffered/mj5.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_5$site<-"Mojave_5"

Panoche_1 <- read_delim("gbif_sites_buffered/pan1.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_1$site<-"Panoche_1"

Panoche_2 <- read_delim("gbif_sites_buffered/pano2.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_2$site<-"Panoche_2"

Panoche_3 <- read_delim("gbif_sites_buffered/pan3.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_3$site<-"Panoche_3"

Panoche_sl <- read_delim("gbif_sites_buffered/pansl.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_sl$site<-"Panoche_sl"

Semitropic <- read_delim("gbif_sites_buffered/semit.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Semitropic$site<-"Semitropic"

Sheephole <- read_delim("gbif_sites_buffered/sheep.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Sheephole$site<-"Sheephole"

Tejon <- read_delim("gbif_sites_buffered/tej.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Tejon$site<-"Tejon"

Tecopa <- read_delim("gbif_sites_buffered/tec.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Tecopa$site<-"Tecopa"

gbif_sites_buffered<-rbind(Avenal,Barstow_2,Carrizo_1,Carrizo_2,Carrizo_3,Carrizo_4,Carrizo_6,Cuyama_1,Cuyama_2,Cuyama_3,Cuyama_4,Cuyama_5,Cuyama_6,HeartofMojave,Lokern,Mojave_1,Mojave_2,Mojave_3,Mojave_4,Mojave_5,Panoche_1,Panoche_2,Panoche_3,Panoche_sl,Semitropic,Sheephole,Tecopa,Tejon)
```

filter out nas, create new rows for individualcount>1
```{r}
#if individual count = na
gbif_sites_buffered<-gbif_sites_buffered %>% drop_na(individualCount)

gbif_buffered_uniques<-gbif_sites_buffered%>%distinct(gbifID, .keep_all = TRUE)

expanded_buffered_uniques<- data.frame(gbif_buffered_uniques[rep(seq_len(dim(gbif_buffered_uniques)[1]),  with(gbif_buffered_uniques, ifelse(individualCount > 0 & !is.na(individualCount), individualCount, 1))
), , drop = FALSE], row.names=NULL)

expanded_buffered_uniques$individualCount<-1

expanded_buffered_uniques<-rename(expanded_buffered_uniques,lat=decimalLatitude)
expanded_buffered_uniques<-rename(expanded_buffered_uniques,long=decimalLongitude)
expanded_buffered_uniques<-rename(expanded_buffered_uniques,site_code=site)

#if individual count >1
expanded_sites <- data.frame(gbif_sites_buffered[rep(seq_len(dim(gbif_sites_buffered)[1]),  with(gbif_sites_buffered, ifelse(individualCount > 0 & !is.na(individualCount), individualCount, 1))
), , drop = FALSE], row.names=NULL)

expanded_sites$individualCount<-1

expanded_sites<-rename(expanded_sites,lat=decimalLatitude)
expanded_sites<-rename(expanded_sites,long=decimalLongitude)
expanded_sites<-rename(expanded_sites,site_code=site)
#there are duplicates due to overlap

write.csv(expanded_sites, file = "buffer_gbif.csv")

write.csv(expanded_buffered_uniques, file = "buffer_gbif.csv")
```