---
title: "Plots"
author: "Zoë Hillier-Weltman"
date: "23/08/2022"
output:
  html_document: default
---
```{r, echo=FALSE}
library(ggplot2)
library(here)
library(tidyverse)
library(scales)
library(sf)
library(units)
library(dplyr)
library(raster)
library(spData)
library(tmap) 
```

load files
```{r, echo=FALSE}
expanded<-readRDS('site_gbif.rds')
expanded_sites<-readRDS('buffered_site_gbif.rds')
expanded_regions<-readRDS('region_gbif.rds')
```

shrub files
```{r, echo=FALSE}
panoche<-read.csv("Panoche.csv")
panoche$region<-("Panoche")
carrizo<-read.csv("Carrizo.csv")%>%
  drop_na()
carrizo$lat <- as.numeric(as.character(carrizo$lat))
cuyama<-read.csv("Cuyama.csv")
cuyama$region<-("Cuyama")
carrizo$region<-("Carrizo")
heartofmojave<-read.csv("Heartofmojave.csv", colClasses=c("site"="character"))
names(heartofmojave)[3]<-paste("site_code") 
tecopa<-read.csv("Tecopa.csv")
sheephole<-read.csv("Sheephole.csv")
barstow1<-read.csv("Barstow_1.csv", colClasses=c("site"="character"))
barstow2<-read.csv("barstow_2.csv", colClasses=c("site"="character"))
avenal<-read.csv("Avenal.csv")
coalinga<-read.csv("Coalinga.csv")
semitropic<-read.csv("Semitropic.csv")
mojave<-read.csv("Mojave.csv" , colClasses=c("site"="character"))
lokern<-read.csv("Lokern.csv")
tejon<-read.csv("Tejon_ERG.csv", colClasses=c("site"="character"))

semitropic$region<-"San Joaquin"
coalinga$region<-"Coalinga"
avenal$region<-"San Joaquin"
lokern$region<-"Lokern"
tejon$region<-"San Joaquin"

new_shrubs<-full_join(panoche,carrizo)
new_shrubs<-full_join(new_shrubs,cuyama)
new_shrubs<-full_join(new_shrubs,tecopa)
new_shrubs<-full_join(new_shrubs,sheephole)
new_shrubs<-full_join(new_shrubs,barstow1)
new_shrubs<-full_join(new_shrubs,barstow2)
new_shrubs<-full_join(new_shrubs,heartofmojave)
new_shrubs<-full_join(new_shrubs,coalinga)
new_shrubs<-full_join(new_shrubs,semitropic)
new_shrubs<-full_join(new_shrubs,avenal)
new_shrubs<-full_join(new_shrubs,mojave)
new_shrubs<-full_join(new_shrubs,lokern)
new_shrubs<-full_join(new_shrubs,tejon)

shrubs_sites<-new_shrubs%>%group_by(site_code,region)%>%summarise()
```

polygons
```{r, echo=FALSE}
data_path <- "sites"

read_zip = function(f){
 s = sf::st_read(f)
 return(s[,c("Name")])
}
polygons_read <- as.list(dir(data_path, pattern="*.kml", full.names=T)) 
polygons<-do.call(rbind, lapply(polygons_read, read_zip))

prog_polys<-polygons%>%
  dplyr::filter(Name %in% (c("Avenal","Barstow_1","Barstow_2","Carrizo_1","Carrizo_2","Carrizo_3","Carrizo_4","Carrizo_6","Coalinga","Cuyama_1","Cuyama_2","Cuyama_3","Cuyama_4","Cuyama_5","Cuyama_6","HeartofMojave","Lokern","Mojave_1","Mojave_2","Mojave_3","Mojave_4","Mojave_5","Panoche_1","Panoche_2","Panoche_3","Panoche_silvercreek","Semitropic","Sheephole","Tejon_ERG","Tecopa"
                )))
```

birds/scales for top ten species of the three scales
```{r,echo=F, echo=FALSE}
#note that expanded is original sites, expanded_sites is buffered sites, expanded_regions is regions

birds_species_2<-expanded_sites%>%
  group_by(month, individualCount, site_code,species)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Summer",
    month < 3  ~ "Winter",
    month >= 9 ~ "Winter"
    ))%>%
  aggregate(individualCount ~season+species, sum)

topten_winter<-birds_species_2%>%group_by(species,season)%>%
  filter(season=="Winter")%>%
    arrange(desc(individualCount))
topten_winter <- topten_winter[1:10,]

topten_summer<-birds_species_2%>%group_by(species,season)%>%
  filter(season=="Summer")%>%
    arrange(desc(individualCount))
topten_summer <- topten_summer[1:10,]

birds_species_seasons<-full_join(topten_summer,topten_winter)


#remove scale as a factor and add season

ggplot(birds_species_seasons, aes(fill=season, y=individualCount, x=species))+
geom_bar(position='dodge', stat='identity')+
  theme_classic()+
    xlab("Species")+
  ylab("Number of bird observations")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))


#nspecies by site by season
birds_species_2<-expanded_sites%>%
  group_by(month, individualCount, site_code,species)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Summer",
    month < 3  ~ "Winter",
    month >= 9 ~ "Winter"
    ))%>%
  aggregate(individualCount ~season+species+site_code, sum)

species_sites<-birds_species_2%>%group_by(species,site_code,season)%>%
 summarise()
species_sites$species_count<-1

species_sites<-species_sites%>%
group_by(site_code,species_count,season) %>%
  aggregate(species_count~site_code+season,sum)

ggplot(species_sites, aes(fill=season, y=species_count, x=site_code))+
geom_bar(position='dodge', stat='identity')+
  theme_classic()+
    xlab("Site")+
  ylab("Number of species")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))

glm<-glm(species_count~site_code+season,data=species_sites)
kruskal.test(x='species_count',g='site_code',data=species_sites)
```
Why is your bird variable the number of bird observations per site?   E.g. cumulative number of time….?

How many sites have bird observations?  


Mean number of birds per site, per season, per year?    Variance in abundance across sites?  
```{r}
expanded_sites%>%
  group_by(site_code)%>%
  summarise()

birds_sites_seasons_year<-expanded_sites%>%
  group_by(month, individualCount, site_code,year)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Summer",
    month < 3  ~ "Winter",
    month >= 9 ~ "Winter"
    ))%>%
  aggregate(individualCount ~season+site_code+year, sum)

#mean n per site per season
sum_site_season<-aggregate(individualCount ~season+site_code, birds_sites_seasons_year, sum)

ggplot(sum_site_season, aes(fill=season, y=individualCount, x=site_code))+
geom_bar(position='dodge', stat='identity')+
  theme_classic()+
    xlab("Site")+
  ylab("Number of bird observations")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))

#mean n per site per year
mean_site_year<-aggregate(individualCount ~year+site_code, birds_sites_seasons_year, mean)

ggplot(mean_site_year, aes(color=site_code, y=individualCount, x=year))+
geom_point()+
  theme_classic()+
    xlab("Year")+
  ylab("Mean number of bird observations")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))

#mean n per season per year

sum_year_season<-aggregate(individualCount ~season+year, birds_sites_seasons_year, mean)

ggplot(sum_year_season, aes(color=season, y=individualCount, x=year))+
  geom_point()+
geom_smooth(se=F)+
  theme_classic()+
    xlab("Year")+
  ylab("Mean number of bird observations")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))

year_season_site<-aggregate(individualCount ~season+year+site_code, birds_sites_seasons_year, sum)

year_count<-aggregate(individualCount ~year, birds_sites_seasons_year, sum)

#var in abundance across sites
library(mvabund)
library(pastecs)
library(gtable)

birddat <- mvabund(expanded_sites$abund)
birddat <- as.mvabund(year_count)
mvabund::meanvar.plot(birddat)

glm<-manyglm(individualCount ~season+year+site_code,data=year_season_site)
glm_stats<-summary.manyglm(glm)
#table(glm_stats)

topten_winter<-birds_species_2%>%group_by(species,season)%>%
  filter(season=="Winter")%>%
    arrange(desc(individualCount))
topten_winter <- topten_winter[1:10,]

topten_summer<-birds_species_2%>%group_by(species,season)%>%
  filter(season=="Summer")%>%
    arrange(desc(individualCount))
topten_summer <- topten_summer[1:10,]

birds_species_seasons<-full_join(topten_summer,topten_winter)

#chisq<-birds_seasons_2%>%
#  subset(select=c(individualCount,season))
#stats::chisq.test(chisq$individualCount,chisq$season)
#glm<-glm(chisq)
#anova(glm)

#remove scale as a factor and add season

ggplot(birds_species_seasons, aes(fill=season, y=individualCount, x=species))+
geom_bar(position='dodge', stat='identity')+
  theme_classic()+
    xlab("Species")+
  ylab("Number of bird observations")+ 
  coord_cartesian(expand=F)+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
```


birds/regions
```{r, echo=FALSE}
gbif_region_year<-expanded_regions%>%
  group_by(year)%>%
  summarise(n=n())

ggplot(gbif_region_year, aes(x=year, y=n))+
  geom_point()+
  geom_smooth(se=F)+
  coord_cartesian(expand=F)+
  theme_classic()+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))+
  xlab("Year")+
  ylab("Number of birds observed")+
  ggtitle("Total Regional Bird Counts per Year")
```

birds by month, buffered sites
```{r, echo=FALSE}
birds_seasons_regions<-expanded_regions%>%
  group_by(month, individualCount)%>%
  aggregate(individualCount ~month, sum)

ggplot(birds_seasons_regions,aes(x=month,y=individualCount))+
  geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
  xlab("Month")+
  ylab("Number of birds observed")+
   scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 12))
```

shrubs and sites
```{r, echo=FALSE}
birds_seasons_sites_2<-expanded_sites%>%
  group_by(month, individualCount)%>%
  aggregate(individualCount ~month, sum)

ggplot(birds_seasons_sites_2,aes(x=month,y=individualCount))+
  geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
  xlab("Month")+
  ylab("Number of birds observed")+
   scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
  scale_x_continuous(breaks = scales::pretty_breaks(n = 12))
```

shrubs/km^2 and n birds

```{r, echo=FALSE}
n_shrubs <- new_shrubs %>% 
  group_by(site_code) %>% 
    dplyr::summarise(n_shrubs = n())

n_shrubs$n_shrubs=as.numeric(levels(n_shrubs$n_shrubs))[n_shrubs$n_shrubs]

polygons$area<-st_area(polygons)
names(polygons)[1]<-paste("site_code") 
polygons<-polygons%>%mutate(site_code = recode(site_code, 'HeartofMojave' = "Heartofmojave") )
shrubs_by_area<-merge(new_shrubs,polygons, by='site_code')

shrubs_by_area<-shrubs_by_area %>% 
  group_by(site_code,area) %>% 
    dplyr::summarise(n_shrubs = n())%>%
  mutate('shrubs_m^2' = n_shrubs/area)

#shrubs/km^2 since m^2 is tiny
shrubs_by_area_km<-shrubs_by_area %>% 
  group_by(site_code,area) %>% 
  mutate(area=area/1000000)%>%
    dplyr::summarise(n_shrubs = n())%>%
  mutate('shrubs_km' = n_shrubs/area)%>%
  drop_units()

#summer

birds_shrubs_2_summer<-expanded_sites%>%
  group_by(month, individualCount, site_code)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Non-Breeding",
    month < 3  ~ "Breeding",
    month >= 9 ~ "Breeding"
    ))%>%
  aggregate(individualCount ~season+site_code, sum)%>%
  filter(season=="Non-Breeding")

birds_shrubs_km_summer<-full_join(birds_shrubs_2_summer,shrubs_by_area_km)%>%
  drop_na()

#winter

birds_shrubs_winter<-expanded_sites%>%
  group_by(month, individualCount, site_code)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Non-Breeding",
    month < 3  ~ "Breeding",
    month >= 9 ~ "Breeding"
    ))%>%
  aggregate(individualCount ~season+site_code, sum)%>%
    filter(season=="Breeding")

birds_shrubs_all<-expanded_sites%>%
  group_by(month, individualCount, site_code)%>%
  summarise(individualCount=n())%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Summer",
    month < 3  ~ "Winter",
    month >= 9 ~ "Winter"
    ))%>%
  aggregate(individualCount ~season+site_code, sum)

birds_shrubs_km_all<-full_join(birds_shrubs_all,shrubs_by_area_km)%>%
  drop_na()

#summer, buffered

ggplot(birds_shrubs_km_summer, aes(y=individualCount, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Bird count")+ 
  ggtitle("Shrub Density and Summer Bird Observations")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#winter, buffered

ggplot(birds_shrubs_km_winter, aes(y=individualCount, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Bird count")+ 
  ggtitle("Shrub Density and Winter Bird Observations")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#both

ggplot(birds_shrubs_km_all, aes(y=individualCount, x=shrubs_km,fill=season,color=season))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Bird count")+ 
  ggtitle("Shrub Density and Bird Observations")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#variance between

glm<-glm(individualCount~shrubs_km+season+site_code, data=birds_shrubs_km_all,family="quasipoisson")
anova(glm)
  print(chisq.test(birds_shrubs_km_all$individualCount))
  
  chisq.test(birds_shrubs_km_all$shrubs_km)
  t.test(birds_shrubs_km_all$individualCount)
    t.test(birds_shrubs_km_all$individualCount)
        t.test(birds_shrubs_km_all$shrubs_km)
        
kruskal.test(individualCount~season, data=birds_shrubs_km_all)   
kruskal.test(shrubs_km~site_code, data=birds_shrubs_km_all) 
kruskal.test(individualCount~site_code, data=birds_shrubs_km_all) 
  #for two factors in a t test, need to be paired, eg seasons
  #so do one factor for count and shrub each, then seasons and count

library(FSA)
DT = dunnTest(shrubs_km~site_code, data=birds_shrubs_km_all)
DT
PT = DT$res
PT

library(rcompanion)

cldList(P.adj ~ Comparison,
        data = PT,
        threshold = 0.05)
dunnTest(individualCount~site_code, data=birds_shrubs_km_all)

#save posthoc for later, once have aridity

glm<-glm(individualCount~shrubs_km+season+site_code, data=birds_shrubs_km_all)
anova(glm)
library(emmeans)
emmeans(glm,~shrubs_km)#not helpful

cor.test(birds_shrubs_km$individualCount,birds_shrubs_km$shrubs_km)
```
shrubs/km^2 and n species

```{r, echo=FALSE}
n_shrubs <- new_shrubs %>% 
  group_by(site_code) %>% 
    dplyr::summarise(n_shrubs = n())

n_shrubs$n_shrubs=as.numeric(levels(n_shrubs$n_shrubs))[n_shrubs$n_shrubs]

polygons$area<-st_area(polygons)
names(polygons)[1]<-paste("site_code") 
polygons<-polygons%>%mutate(site_code = recode(site_code, 'HeartofMojave' = "Heartofmojave") )
shrubs_by_area<-merge(new_shrubs,polygons, by='site_code')

shrubs_by_area<-shrubs_by_area %>% 
  group_by(site_code,area) %>% 
    dplyr::summarise(n_shrubs = n())%>%
  mutate('shrubs_m^2' = n_shrubs/area)

#shrubs/km^2 since m^2 is tiny
shrubs_by_area_km<-shrubs_by_area %>% 
  group_by(site_code,area) %>% 
  mutate(area=area/1000000)%>%
    dplyr::summarise(n_shrubs = n())%>%
  mutate('shrubs_km' = n_shrubs/area)%>%
  drop_units()

#summer

species_shrubs_2_summer<-expanded_sites%>%
  group_by(month, species, site_code)%>%
  summarise()%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Non-Breeding",
    month < 3  ~ "Breeding",
    month >= 9 ~ "Breeding"
    ))%>%
    filter(season=="Non-Breeding")

species_sites_summer<-species_shrubs_2_summer%>%group_by(species,site_code,season)%>%
 summarise()
species_sites_summer$species_count<-1

species_sites_summer<-species_sites_summer%>%
group_by(site_code,species_count,season) %>%
  aggregate(species_count~site_code,sum)


  #aggregate(species ~season+site_code, sum)%>%

species_shrubs_km_summer<-full_join(species_sites_summer,shrubs_by_area_km)%>%
  drop_na()

#winter

species_shrubs_2_winter<-expanded_sites%>%
  group_by(month, species, site_code)%>%
  summarise()%>%
    mutate(season = case_when(
    month >= 3 & month <9 ~ "Non-Breeding",
    month < 3  ~ "Breeding",
    month >= 9 ~ "Breeding"
    ))%>%
    filter(season=="Breeding")

species_sites_winter<-species_shrubs_2_winter%>%group_by(species,site_code,season)%>%
 summarise()
species_sites_winter$species_count<-1

species_sites_winter<-species_sites_winter%>%
group_by(site_code,species_count,season) %>%
  aggregate(species_count~site_code,sum)

species_shrubs_km_winter<-full_join(species_sites_winter,shrubs_by_area_km)%>%
  drop_na()

###combined

cor.test(species_shrubs_km$species_count,species_shrubs_km$shrubs_km)

species_sites<-species_shrubs_2_winter%>%group_by(species,site_code,season)%>%
 summarise()
species_sites_winter$species_count<-1

species_sites_winter<-species_sites_winter%>%
group_by(site_code,species_count,season) %>%
  aggregate(species_count~site_code,sum)

species_shrubs_km<-full_join(species_sites,shrubs_by_area_km)%>%
  drop_na()

#summer, buffered

ggplot(birds_shrubs_km_summer, aes(y=individualCount, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Bird count")+ 
  ggtitle("Shrub Density and Summer Bird Observations")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#winter, buffered

ggplot(birds_shrubs_km_winter, aes(y=individualCount, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Bird count")+ 
  ggtitle("Shrub Density and Winter Bird Observations")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#for species

ggplot(species_shrubs_km_summer, aes(y=species_count, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Species count")+ 
  ggtitle("Shrub Density and Summer Species Count")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

#winter, buffered

ggplot(species_shrubs_km_winter, aes(y=species_count, x=shrubs_km))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Species count")+ 
  ggtitle("Shrub Density and Winter Species Count")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

ggplot(species_shrubs_km, aes(y=species_count, x=shrubs_km,fill=season,color=season))+
geom_point()+
  theme_classic()+
  geom_smooth(se=F)+
    xlab("Shrubs per square kilometer")+
  ylab("Species count")+ 
  ggtitle("Shrub Density and Species Count")+
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) 

```


shrub density and sites
```{r, echo=FALSE}
ggplot(shrubs_by_area_km,aes(x=site_code,y=shrubs_km))+
  theme_classic()+
  geom_col()+
  coord_cartesian(expand=T)+
  xlab("Site")+
  ylab("Shrubs per km^2")+ 
  scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 20))+
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
```
number of species/site
```{r, echo=FALSE}
n_species_sites<-expanded_sites%>%group_by(species,site_code,individualCount)%>%
  aggregate(individualCount~species+site_code,sum)%>%
arrange(desc(species)) 
n_species_sites$species_count<-1

n_species_sites<-n_species_sites%>%
group_by(site_code,species_count,species) %>%
  aggregate(species_count~site_code,sum)

ggplot(n_species_sites,aes(x=site_code,y=species_count))+
  geom_bar(position='dodge', stat='identity')+
  theme_classic()+
  geom_smooth(se=F)+
  xlab("Site")+
  ylab("Number of species")+
   scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
```
site/number of birds-buffered
```{r}
n_birds_sites<-expanded_sites%>%group_by(site_code,individualCount)%>%
  aggregate(individualCount~site_code,sum)

ggplot(n_birds_sites,aes(x=site_code,y=individualCount))+
  geom_bar(position='dodge', stat='identity')+
  theme_classic()+
  geom_smooth(se=F)+
  xlab("Site")+
  ylab("Number of bird observations")+
   scale_y_continuous(labels=comma,breaks = scales::pretty_breaks(n = 10))+
    theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
```

table formatting is wonky

tables of species-regions
```{r, echo=FALSE}
topten_5<-expanded_regions%>%group_by(species,region)%>%
  aggregate(individualCount~species+region,sum)%>%
arrange(desc(individualCount)) %>% 
group_by(region) %>% slice(1:10)

knitr::kable(topten_5, format="html")
```

table of species-sites
```{r, echo=FALSE}
topten_4<-expanded_sites%>%group_by(species,site_code)%>%
  aggregate(individualCount~species+site_code,sum)%>%
arrange(desc(individualCount)) %>% 
group_by(site_code) %>% slice(1:10)

knitr::kable(topten_4, format="html")
```

map of plots
```{r}

plot(st_geometry(polygons))

library(ggmap)
caMap = map_data("state")%>%
  filter(region =="california")

library(raster)

###
library(tidyr)
library(broom)

shpfile <- prog_polys%>%
  st_zm()
  shpfile<-as_Spatial(shpfile)
  shpfile<-spTransform(shpfile,"+init=epsg:4326") # transform coordinates
tidydta2 <- tidy(shpfile, group=group) 

library(rgeos)
tidydta <- gCentroid(shpfile, byid = TRUE)

library(plyr)
# Recover row name 
temp_df <- data.frame(prog_polys$Name)
names(temp_df) <- c("region")
# Create and append "id"
temp_df$id <- seq(1,nrow(temp_df))

new_df <- join(tidydta2, temp_df)
df<-new_df[!duplicated(new_df$region),]
df-as_Spatial(df)
df<-spTransform(df,"+init=epsg:4326")


library(ggrepel)

ggplot(data = caMap) + 
  geom_polygon(aes(x=long,y=lat),color = "black", fill = "gray" ) +
  coord_map(xlim = c(-122.5, -115.0),ylim = c(32.5,37.5))+
  geom_point(aes(x=long,y=lat,label=region),
               data=df,
               size=2)+
geom_text_repel()



plot(buffered_centroids)

ggplot(data = caMap) + 
  geom_polygon(aes(x=long,y=lat),color = "black", fill = "gray" ) +
  coord_map(xlim = c(-122.5, -115.0),ylim = c(32.5,37.5))+
  geom_point(data=buffered_centroids)+
geom_text_repel()

ggplot(data = caMap) + 
  geom_polygon(aes(x=long,y=lat),color = "black", fill = "gray" ) +
  coord_map(xlim = c(-122.5, -115.0),ylim = c(32.5,37.5))+
  geom_point(aes(x=X,y=Y,fill=Name),data=sites,size = 2,shape=23)+
  xlab("Longitude (degrees)")+
  ylab("Latitude (degrees)")+
  ggtitle("Map of Sites")+ 
  theme(legend.title = element_text(size = 6), 
               legend.text = element_text(size = 5))+
  theme_classic()
        
sites<-cbind(centroids_prog, st_coordinates(centroids_prog))

```

