---
title: "occupancy models"
author: "Zoë Hillier-Weltman"
date: "23/09/2022"
output: html_document
---
from ebird
```{r}
library(sf)
library(terra)
library(raster)
library(rgdal)
library(MODIS)
library(exactextractr)
library(viridis)
library(tidyverse)
# resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection

#ebird example
# bcr 27 boundary
bcr <- read_sf("ebird data/gis-data.gpkg", "bcr") %>% 
  filter(bcr_code == 27) %>% 
  # project to the native modis projection
  st_transform(crs = paste("+proj=sinu +lon_0=0 +x_0=0 +y_0=0",
                           "+a=6371007.181 +b=6371007.181 +units=m +no_defs"))
# load ebird data
ebird <- readr::read_csv("ebird data/ebd_woothr_june_bcr27_zf.csv")
# get list of tiles required to cover this bcr
tiles <- getTile(bcr)
tiles@tile
#> [1] "h10v06" "h10v05" "h11v05"
```

for gbif
```{r}
#modified for gbif data
# bcr 27 boundary
bcr <- read_sf("data/gis-data.gpkg", "bcr") %>% 
  filter(bcr_code == 27) %>% 
  # project to the native modis projection
  st_transform(crs = paste("+proj=sinu +lon_0=0 +x_0=0 +y_0=0",
                           "+a=6371007.181 +b=6371007.181 +units=m +no_defs"))
# load ebird data
ebird <- read_csv("data/ebd_woothr_june_bcr27_zf.csv")
# get list of tiles required to cover this bcr
tiles <- getTile(bcr)
tiles@tile
#> [1] "h10v06" "h10v05" "h11v05"
```

ebird
```{r}
#library(auk)
MODIS::EarthdataLogin(usr = "zoehw", pwd = "Pumpkinman1!")

# earliest year of ebird data
begin_year <- format(min(ebird$observation_date), "%Y.01.01")
# end date for ebird data
end_year <- format(max(ebird$observation_date), "%Y.12.31")
# download tiles and combine into a single raster for each year
tifs <- runGdal(product = "MCD12Q1", collection = "006", SDSstring = "01", 
                extent = bcr %>% st_buffer(dist = 10000), 
                begin = begin_year, end = end_year, 
                outDirPath = "data", job = "modis",
                MODISserverOrder = "LPDAAC") %>% 
  pluck("MCD12Q1.006") %>% 
  unlist()

# rename tifs to have more descriptive names
new_names <- format(as.Date(names(tifs)), "%Y") %>% 
  sprintf("modis_mcd12q1_umd_%s.tif", .) %>% 
  file.path(dirname(tifs), .)
file.rename(tifs, new_names)
```

```{r}
# load the landcover data
landcover <- list.files("data/modis", "^modis_mcd12q1_umd", 
                        full.names = TRUE) %>% 
  stack()
# label layers with year
landcover <- names(landcover) %>% 
  str_extract("(?<=modis_mcd12q1_umd_)[0-9]{4}") %>% 
  paste0("y", .) %>% 
  setNames(landcover, .)
landcover
#> class      : RasterStack 
#> dimensions : 1969, 4301, 8468669, 9  (nrow, ncol, ncell, nlayers)
#> resolution : 463, 463  (x, y)
#> extent     : -8745491, -6752783, 3242262, 4154525  (xmin, xmax, ymin, ymax)
#> crs        : +proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs 
#> names      : y2010, y2011, y2012, y2013, y2014, y2015, y2016, y2017, y2018 
#> min values :     0,     0,     0,     0,     0,     0,     0,     0,     0 
#> max values :   255,   255,   255,   255,   255,   255,   255,   255,   255
```
names are printing yNA.1, etc instead of y2010 etc
so max integer below is non-numeric
fix

```{r}
max_lc_year <- names(landcover) %>% 
  str_extract("[0-9]{4}") %>% 
  as.integer() %>% 
  max()
```
modis data now loaded for these specs

landscape metrics- ebird data matched to modis
note likely won't yield anything until i fix year issue
```{r}
neighborhood_radius <- 5 * ceiling(max(res(landcover))) / 2
ebird_buff <- ebird %>% 
  distinct(year = format(observation_date, "%Y"),
           locality_id, latitude, longitude) %>% 
  # for 2019 use 2018 landcover data
  mutate(year_lc = if_else(as.integer(year) > max_lc_year, 
                           as.character(max_lc_year), year),
         year_lc = paste0("y", year_lc)) %>% 
  # convert to spatial features
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  # transform to modis projection
  st_transform(crs = projection(landcover)) %>% 
  # buffer to create neighborhood around each point
  st_buffer(dist = neighborhood_radius) %>% 
  # nest by year
  nest(data = c(year, locality_id, geometry))
```

```{r}
# function to summarize landcover data for all checklists in a given year
calculate_pland <- function(yr, regions, lc) {
  locs <- st_set_geometry(regions, NULL)
  exact_extract(lc[[yr]], regions, progress = FALSE) %>% 
    map(~ count(., landcover = value)) %>% 
    tibble(locs, data = .) %>% 
    unnest(data)
}
# iterate over all years extracting landcover for all checklists in each
lc_extract <- ebird_buff %>% 
  mutate(pland = map2(year_lc, data, calculate_pland, lc = landcover)) %>% 
  select(pland) %>% 
  unnest(cols = pland)
```

proportion of neighbourhood within each land cover class
```{r}
pland <- lc_extract %>% 
  # calculate proporiton
  group_by(locality_id, year) %>% 
  mutate(pland = n / sum(n)) %>% 
  ungroup() %>% 
  select(-n) %>% 
  # remove NAs after tallying so pland is relative to total number of cells
  filter(!is.na(landcover))
```

desc names
```{r}
# convert names to be more descriptive
lc_names <- tibble(landcover = 0:15,
                   lc_name = c("pland_00_water", 
                               "pland_01_evergreen_needleleaf", 
                               "pland_02_evergreen_broadleaf", 
                               "pland_03_deciduous_needleleaf", 
                               "pland_04_deciduous_broadleaf", 
                               "pland_05_mixed_forest",
                               "pland_06_closed_shrubland", 
                               "pland_07_open_shrubland", 
                               "pland_08_woody_savanna", 
                               "pland_09_savanna", 
                               "pland_10_grassland", 
                               "pland_11_wetland", 
                               "pland_12_cropland", 
                               "pland_13_urban", 
                               "pland_14_mosiac", 
                               "pland_15_barren"))
pland <- pland %>% 
  inner_join(lc_names, by = "landcover") %>% 
  arrange(landcover) %>% 
  select(-landcover)

# tranform to wide format, filling in implicit missing values with 0s%>% 
pland <- pland %>% 
  pivot_wider(names_from = lc_name, 
              values_from = pland, 
              values_fill = list(pland = 0))

# save
write_csv(pland, "data/modis_pland_location-year.csv")
```

habitat covars
```{r}
agg_factor <- round(2 * neighborhood_radius / res(landcover))
r <- raster(landcover) %>% 
  aggregate(agg_factor) 
r <- bcr %>% 
  st_transform(crs = projection(r)) %>% 
  rasterize(r, field = 1) %>% 
  # remove any empty cells at edges
  trim()
r <- writeRaster(r, filename = "data/prediction-surface.tif", overwrite = TRUE)
```

pland again
```{r}
# get cell centers and create neighborhoods
r_centers <- rasterToPoints(r, spatial = TRUE) %>% 
  st_as_sf() %>% 
  transmute(id = row_number())
r_cells <- st_buffer(r_centers, dist = neighborhood_radius)

# extract landcover values within neighborhoods, only needed most recent year
lc_extract_pred <- landcover[[paste0("y", max_lc_year)]] %>% 
  exact_extract(r_cells, progress = FALSE) %>% 
  map(~ count(., landcover = value)) %>% 
  tibble(id = r_cells$id, data = .) %>% 
  unnest(data)

# calculate the percent for each landcover class
pland_pred <- lc_extract_pred %>% 
  count(id, landcover) %>% 
  group_by(id) %>% 
  mutate(pland = n / sum(n)) %>% 
  ungroup() %>% 
  select(-n) %>% 
  # remove NAs after tallying so pland is relative to total number of cells
  filter(!is.na(landcover))

# convert names to be more descriptive
pland_pred <- pland_pred %>% 
  inner_join(lc_names, by = "landcover") %>% 
  arrange(landcover) %>% 
  select(-landcover)

# tranform to wide format, filling in implicit missing values with 0s
pland_pred <- pland_pred %>% 
  pivot_wider(names_from = lc_name, 
              values_from = pland, 
              values_fill = list(pland = 0)) %>% 
  mutate(year = max_lc_year) %>% 
  select(id, year, everything())

# join in coordinates
pland_coords <- st_transform(r_centers, crs = 4326) %>% 
  st_coordinates() %>% 
  as.data.frame() %>% 
  cbind(id = r_centers$id, .) %>% 
  rename(longitude = X, latitude = Y) %>% 
  inner_join(pland_pred, by = "id")
```

spatial format to map
```{r}
forest_cover <- pland_coords %>% 
  # convert to spatial features
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(crs = projection(r)) %>% 
  # rasterize points
  rasterize(r, field = "pland_04_deciduous_broadleaf") %>% 
  # project to albers equal-area for mapping
  projectRaster(crs = st_crs(102003)$proj4string, method = "ngb") %>% 
  # trim off empty edges of raster
  trim()

# make a map
par(mar = c(0.25, 0.25, 2, 0.25))
t <- str_glue("Proportion of Deciduous Broadleaf Forest\n",
              "{max_lc_year} MODIS Landcover")
plot(forest_cover, axes = FALSE, box = FALSE, col = viridis(10), main = t)
```

elevation etc need to be done manually
```{r}

```

load, crop to modis size, reproject
```{r}
elev <- raster("data/elevation_1KMmd_GMTEDmd.tif")
# crop, buffer bcr by 10 km to provide a little wiggly room
elev <- bcr %>% 
  st_buffer(dist = 10000) %>% 
  st_transform(crs = projection(elev)) %>% 
  crop(elev, .) %>% 
  projectRaster(crs = projection(landcover))
```

for neighbourhoods
```{r}
# buffer each checklist location
ebird_buff_noyear <- ebird %>% 
  distinct(locality_id, latitude, longitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(crs = projection(elev)) %>% 
  st_buffer(dist = neighborhood_radius)

# extract elevation values and calculate median and sd
locs <- st_set_geometry(ebird_buff_noyear, NULL) %>% 
    mutate(id = row_number())
elev_checklists <- exact_extract(elev, ebird_buff_noyear, progress = FALSE) %>% 
  map_dfr(~ tibble(elevation_median = mean(.$value, na.rm = TRUE),
                   elevation_sd = sd(.$value, na.rm = TRUE))) %>% 
  # join to lookup table to get locality_id
  bind_cols(locs, .)
```

for prediction surface
```{r}
# extract and calculate median and sd
elev_pred <- exact_extract(elev, r_cells, progress = FALSE) %>% 
  map_dfr(~ tibble(elevation_median = mean(.$value, na.rm = TRUE),
                   elevation_sd = sd(.$value, na.rm = TRUE))) %>% 
  # join to lookup table to get locality_id
  bind_cols(st_drop_geometry(r_cells), .)
```

combine covars
```{r}
# checklist covariates
pland_elev_checklist <- inner_join(pland, elev_checklists, by = "locality_id")
write_csv(pland_elev_checklist, "data/pland-elev_location-year.csv")

# prediction surface covariates
pland_elev_pred <- inner_join(pland_coords, elev_pred, by = "id")
write_csv(pland_elev_pred, "data/pland-elev_prediction-surface.csv")
glimpse(pland_elev_pred)
#> Rows: 90,949
#> Columns: 22
#> $ id                            <int> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25…
#> $ longitude                     <dbl> -77.3, -77.3, -77.3, -77.3, -77.3, -77.2, -77.4, -77.3, -77.3, -77.3, -77.3, -77.2, -77.2…
#> $ latitude                      <dbl> 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2, 37.2,…
#> $ year                          <int> 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018, 2018,…
#> $ pland_00_water                <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_01_evergreen_needleleaf <dbl> 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0…
#> $ pland_02_evergreen_broadleaf  <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_03_deciduous_needleleaf <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_04_deciduous_broadleaf  <dbl> 0.0000, 0.1724, 0.0345, 0.0000, 0.0000, 0.0345, 0.0000, 0.0000, 0.0000, 0.0000, 0.0345, 0…
#> $ pland_05_mixed_forest         <dbl> 0.0000, 0.0000, 0.0000, 0.1379, 0.1379, 0.2759, 0.0000, 0.0000, 0.0345, 0.1379, 0.0345, 0…
#> $ pland_06_closed_shrubland     <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_07_open_shrubland       <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_08_woody_savanna        <dbl> 0.966, 0.828, 0.724, 0.690, 0.862, 0.690, 0.621, 0.966, 0.793, 0.828, 0.793, 0.310, 0.552…
#> $ pland_09_savanna              <dbl> 0.0000, 0.0000, 0.2069, 0.1724, 0.0000, 0.0000, 0.0345, 0.0345, 0.1379, 0.0345, 0.1379, 0…
#> $ pland_10_grassland            <dbl> 0.0345, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0345, 0.0000, 0.0000, 0…
#> $ pland_11_wetland              <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ pland_12_cropland             <dbl> 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0.0000, 0…
#> $ pland_13_urban                <dbl> 0.0000, 0.0000, 0.0345, 0.0000, 0.0000, 0.0000, 0.3448, 0.0000, 0.0000, 0.0000, 0.0000, 0…
#> $ pland_14_mosiac               <dbl> 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000…
#> $ pland_15_barren               <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,…
#> $ elevation_median              <dbl> 40.0, 41.8, 42.6, 42.8, 42.5, 42.8, 45.6, 44.7, 43.6, 41.2, 40.6, 40.9, 42.8, 78.0, 68.1,…
#> $ elevation_sd                  <dbl> 3.75, 2.78, 2.41, 3.28, 1.40, 1.66, 2.74, 3.33, 2.29, 2.32, 2.52, 3.41, 1.53, 8.61, 7.02,…
```

data prep done, now to model species distributions

may need to use random forest models to account for landscape stuff before going on to occupancy models for bird species/counts in particular

#how much of my data is ebird?
```{r}
#all except 3 points!!! from buffered data
```

