---
title: "regional_gbif"
author: "Zoë Hillier-Weltman"
date: "08/08/2022"
output: html_document
---

```{r}
library(tidyverse)
library(sf)
library(sp)
library(rgdal)
library(raster)
library(rgeos)
library(here)
```

recursive
```{r}
data_path <- "sites"

read_zip = function(f){
 s = sf::st_read(f)
 return(s[,c("Name")])
}
polygons_read <- as.list(dir(data_path, pattern="*.kml", full.names=T)) 
polygons<-do.call(rbind, lapply(polygons_read, read_zip))

#changing format
polygon_wkts<-st_as_text(polygons$geometry)

mean(st_area(polygons))#10km^2 should be bigger than 0.27km^2

```

for prog report- 30 initial sites (21 w shrubs)
```{r}
prog_polys<-polygons%>%
  dplyr::filter(Name %in% (c("Avenal","Barstow_1","Barstow_2","Carrizo_1","Carrizo_2","Carrizo_3","Carrizo_4","Carrizo_6","Coalinga","Cuyama_1","Cuyama_2","Cuyama_3","Cuyama_4","Cuyama_5","Cuyama_6","HeartofMojave","Lokern","Mojave_1","Mojave_2","Mojave_3","Mojave_4","Mojave_5","Panoche_1","Panoche_2","Panoche_3","Panoche_silvercreek","Semitropic","Sheephole","Tejon_ERG","Tecopa"
                )))

```

```{r}
points_centroids_prog<-st_centroid(prog_polys)

```

ignore code fragments here
regional for prog
```{r}
#293.161m is radius of avg area
#adding a 100km buffer 
#343.1831m is radius of buffered points
z = 5630 #buffer is in m

#data, read in, set as sf, list projection####


#buffered_regional_centroids_prog<-st_buffer(regional_centroids_prog, z)

regional_centroids_prog<-st_as_sf(regional_centroids_prog)

test <- mapply(function(x) st_point((x)),regional_centroids_prog$x,SIMPLIFY = FALSE)


#trying squares

#make recursive eventually
bboxes_prog = cbind(st_sf(geometry=buffered_regional_centroids_prog$geometry),do.call(rbind,lapply(buffered_regional_centroids_prog$geometry, st_bbox)))

st_area(bboxes_prog$geometry)
st_area(polygons$geometry)


wkts<-mapply( buffered_regional_centroids_prog, FUN=st_as_text)%>%
  st_as_sf()

df<-as.data.frame(wkts)
plot_sf(df)
st_area(df$x)

fun<-function(st_as_text)
st_as_text(buffered_regional_centroids_prog)

wkts=
  cbind(minx=xmin, miny=ymin, maxx=xmax,maxy=ymax)do.call(rbind,lapply(bboxes, bbox2wkt))
```

code for square sites
```{r}
points_centroids_prog$region<-c("San Joaquin","Barstow","Barstow","Carrizo","Carrizo","Carrizo","Carrizo","Carrizo","Coalinga","Cuyama","Cuyama","Cuyama","Cuyama","Cuyama","Cuyama","Heartofmojave","Lokern", "Mojave","Mojave","Mojave","Mojave","Mojave" ,"Panoche","Panoche","Panoche","Panoche","San Joaquin","Sheephole","Tecopa","San Joaquin")

regional_centroids_prog<-points_centroids_prog%>%
  group_by(region)%>%
  summarise(geometry=st_union(geometry))%>%
  st_centroid(geometry)%>%
  st_make_valid()

plot(regional_centroids_prog)

#trying a thing
mojave_sites<-expanded_sites%>%
  group_by(site_code)%>%
  subset(site_code == c( "Sheephole","Heartofmojave","Mojave_1","Mojave_2","Mojave_3","Mojave_4","Mojave_5"))

mojave_birds<-mojave_sites%>%
  group_by(site_code)%>%
  summarise(n=n())

#try throwing coords for the six sites into gbif, see if any overlap/are near, and use that as a centroid
#mojave 1-5 near each other, sheephole and HoM far

#st_combine yields same results as st_union
regional_centroids_prog_2<-points_centroids_prog%>%
  group_by(region)%>%
  summarise(geometry=st_combine(geometry))%>%
  st_centroid(geometry)%>%
  st_make_valid()

library(wellknown)
library(spocc)
library(buffeRs)

x = 4326 #crs for first spatial df

buffered_regional_centroids_prog <- regional_centroids_prog %>% 
  st_as_sf(coords=c("long","lat"), crs = x, remove=FALSE) %>% 
  st_transform(32610)%>%
  st_make_valid()#crs = 32610 is UTM and will return distances in meters

#create buffer for point data

#regional_centroids_prog<-st_as_sf(regional_centroids_prog)

#change projection to meters before running
```

```{r}

bar = sf::st_point(c(1010236.58888633, 3893762.96328113))
bar = sf::st_sfc(bar)
bar = sf::st_sf(bar)
bar<-buffer_square(bar,10000,0)
st_area(bar)
bar<-bar%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

carr = sf::st_point(c(800005.343531313, 3898368.11138416))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

st_area(carr)

coal = sf::st_point(c(742418.338156988, 4010965.2577555))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

cuy = sf::st_point(c(808757.762586225, 3872412.85095825))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

lok = sf::st_point(c(810421.95871811, 3917631.70982213))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

mjv = sf::st_point(c(1149642.43797103, 3918302.18948119))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert()

pan = sf::st_point(c(698801.049981245, 4060946.99265288))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)
  
sj = sf::st_point(c(823247.321006813, 3938053.76677432))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)

tec = sf::st_point(c(1115892.95312855, 3989096.65328527))%>%
  st_sfc()%>%
  st_sf()%>%
  buffer_square(10000,0)%>%
  st_as_sf(crs=32610)%>%
  st_transform(4326)%>%
  st_make_valid()%>%
  sf_convert(group=F)
  
st_area(tec)

regional_wkts_prog<-rbind(bar,carr,coal,cuy,lok,mjv,pan,sj,tec)

```

#looks like there was an issue converting between circular buffers and square boundingboxes
redoing pull for buffered centroids
do this off good wifi
```{r}
library(rgbif)
library(spocc)

user<- "zoehw2"
email <- "zoe.hillier.weltman@gmail.com"
pwd <- "foxhunter"

#aves taxon key is 212

#not sure if there's a better workaround
for (i in regional_wkts_prog[1:3]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in regional_wkts_prog[4:6]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

for (i in regional_wkts_prog[7:9]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}


library(wicket)
validate_wkt(wkts)
wkts<-wkt_reverse(wkts)
wkts<-wkt_correct(wkts)


plot_sf(wkts)

#rerunning for mojave
for (i in regional_wkts_prog[6]){
  
occ_download(
  pred_within(i),
  pred_in("taxonKey", 212),
  pred("hasCoordinate", T),
  format = "SIMPLE_CSV",
  user=user,pwd=pwd,email=email
)
}

```

```{r}
data_path_2 <- "region"

read_csv = function(f){
 s = read_csv(f)
 return()
}

#gbif_read <- as.list(dir(data_path_2, pattern="*.csv", full.names=T)) 
#gbif_buffered<-do.call(rbind, lapply(gbif_read, read_csv))

library(readr)
Carrizo_r <- read_delim("region/carrizo.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Carrizo_r$region<-"Carrizo"

Cuyama_r <- read_delim("region/cuyama.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Cuyama_r$region<-"Cuyama"

Panoche_r <- read_delim("region/panoche.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Panoche_r$region<-"Panoche"

Barstow_r <- read_delim("region/barstow.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Barstow_r$region<-"Barstow"

Coalinga_r <- read_delim("region/coal.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Coalinga_r$region<-"Coalinga"

Lokern_r <- read_delim("region/lokern.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Lokern_r$region<-"Lokern"

SanJoaquin_r <- read_delim("region/sj.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
SanJoaquin_r$region<-"San Joaquin"

Tecopa_r <- read_delim("region/tecopa.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Tecopa_r$region<-"Tecopa"

Mojave_r <- read_delim("region/mojave.csv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
Mojave_r$region<-"Mojave"

regional_gbif<-rbind(Carrizo_r,Cuyama_r,Panoche_r, Tecopa_r,SanJoaquin_r,Lokern_r,Coalinga_r,Barstow_r,Mojave_r)

```

filter out nas, create new rows for individualcount>1
```{r}
#if individual count = na
regional_gbif<-regional_gbif %>% drop_na(individualCount)

#if individual count >1
expanded_regions <- data.frame(regional_gbif[rep(seq_len(dim(regional_gbif)[1]),  with(regional_gbif, ifelse(individualCount > 0 & !is.na(individualCount), individualCount, 1))
), , drop = FALSE], row.names=NULL)

expanded_regions$individualCount<-1

expanded_regions<-rename(expanded_regions,lat=decimalLatitude)
expanded_regions<-rename(expanded_regions,long=decimalLongitude)
expanded_regions$rep<-1:nrow(expanded_regions)

expanded_regions<-expanded_regions[colnames(expanded_regions)[c(52,1:51)]]
 
region_names<-expanded_regions%>%
  group_by(region)%>%
  summarise()

write.csv(expanded_regions, file = "region_gbif.csv")
saveRDS(expanded_regions, file = "region_gbif.rds")
region_gbif<-readRDS('region_gbif.rds')
```

spinning up some mojave data (mojave, kelso,sheephole, heartofmojave)
```{r}

```

